A Journey from Wild to Textbook Data to Reproducibly Refresh the Wages Data from the National Longitudinal Survey of Youth Database

### 2022-02-09 paper/paper.Rmd
Added the explanation of work experience in Section 4.

From the database, we will refresh this data by re-creating the full data with records from survey years 1979 through to 2018 (the most recent year published). We also plan to make some modifications of the variables. For example, we create a categorical race variable instead of the binary variable "black" and "hispanic".  We also include additional variables so that more interesting data exploration can be included in teaching examples: year of survey, age of individual in 1979, year when the highest grade was completed, number of jobs that the individual had in the corresponding year of survey, total number of hours the individual usually works per week, the year when individual started to work, and the number of years the individual worked.

@@

<!--AE comment: In Section 4, you talk about the variables in the original data. “One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection.” I’m not sure what this means. Please clarify.-->
<!--AE comment: “experience in the workforce” should have the period inside the quotations.-->
There are two aspects of the original data that make a direct comparison difficult. The first one is that The time variable provided is "experience in the workforce". It is not clear how this is calculated. One would expect that there is a record of the day the individual first started a job, and this is used to adjust the year of collection. Secondly, wages were inflation-adjusted to 1990 prices, which is not done for the refreshed data.

```{r compare_xp_to_yrworkforce, fig.cap = "The association between the lenght of time in the workforce and the work experience. The color represents the respondents. Generally, the longer someone were in the workforce, the higher the work experience. However, some of the respondents have much lower working experience compared their duration in the workforce.", eval=FALSE}
```

```{r compare_xp_sw}
```

<!--AE comment: “The treatment of unlikely wages differ in the refreshed data.” I think this means “The treatment of anomalous wages differs between the original and refreshed data sets.” Then you say “We opted to use the weights from a robust linear regression to determine what should be set as missing value as described in Section 3.2.1.” I don’t think you were setting missing values though, were you? It was used to determine which values should be imputed, correct? Either way, please clarify.-->
The treatment of unlikely wages differ in the refreshed data. In the original data by @SingerJudithD2003Alda, wages greater than $75 are set to be missing. However, in recent context, this value is too low to be set as the maximum threshold. We opted to use the weights from a robust linear regression to determine what should be regarded as outliers and imputed them with their predicted values as as described in Section \@ref(censor).

>>>>

From the database, we will refresh this data by re-creating the full data with records from survey years 1979 through to 2018 (the most recent year published). We also plan to make some modifications of the variables. For example, we create a categorical race variable instead of the binary variable "Black" and "Hispanic".  We also include additional variables so that more interesting data exploration can be included in teaching examples: year of survey, age of individual in 1979, year when the highest grade was completed, number of jobs that the individual had in the corresponding year of survey, total number of hours the individual usually works per week, the year when individual started to work, and the number of years the individual worked.

@@

There are two aspects of the original data that make the replication with exactly the same result is not obtained. The first one is that the calculation of the time variable provided, which is the "work experience", is not clearly articulated. @SingerJudithD2003Alda mention that the temporal variable is the years of experience since the first day of work. However, this exact same variable is not explicitly available in the database. From the NLSY79 topical guide [@nlsy79guide], we find that several variables are tagged as work experience-related variables. One of them is the weeks of worked since the last interview. Hence, we use this to calculate the variable. Even though we got different result, the correlation between the original and refreshed work experience is reasonably high as shown in Figure \ref{fig:compare-xp-sw}.

```{r compare-xp-sw, fig.cap = "Males dropouts' work experience in the original and refreshed data. Each point represents the work experience of individuals (identified by the ID) in the corresponding survey period. We can see that the correlation is reasonably high between them even though the exactly same figure could not be obtained.", out.width = "70%"}
```

Secondly, in the original dataset, wages were inflation-adjusted to 1990 prices. This is not done for the refreshed data because we plan to keep refreshing it as the data from the newest survey period release. The 1990 prices might not be relevant anymore for the future context. Hence, we keep it as the original value to allow the users to treat it as their needs.

```{r compare-xp-to-yrworkforce, fig.cap = "The association between the lenght of time in the workforce and the work experience. Generally, the longer someone were in the workforce, the higher the work experience. However, some of the respondents have much lower working experience compared to their duration in the workforce. This is possible when someone ", eval=FALSE}
```

<!--AE comment: “The treatment of unlikely wages differ in the refreshed data.” I think this means “The treatment of anomalous wages differs between the original and refreshed data sets.” Then you say “We opted to use the weights from a robust linear regression to determine what should be set as missing value as described in Section 3.2.1.” I don’t think you were setting missing values though, were you? It was used to determine which values should be imputed, correct? Either way, please clarify.-->
It is also worth noting that the treatment of unlikely wages differ in the refreshed data. In the original data by @SingerJudithD2003Alda, wages greater than $75 are set to be missing. However, in recent context, this value is too low to be set as the maximum threshold. We opted to use the weights from a robust linear regression to determine what should be regarded as outliers and imputed them with their predicted values as as described in Section \@ref(censor).
